---
title: "Homework 5/Midterm"
author: "Bryan Hee"
date: "March 3, 2018"
output:
  html_document:
    toc: true
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
```

#The Tidyverse Packages
##Packages
The following list represents tasks which can be accomplished using the tidyverse library and the corresponding package that is associated with each task:

a. Plotting - ggplot2()

b. Data munging/wrangling - dplyr()

c. Reshaping (spreading and gathering) data - tidyr()

d. Importing/exporting data - readr()

##Functions
The following are 2 functions that we've used from each package listed above:

a. ggplot2() - geom_point(), coord_flip()

b. dplyr() - arrange(), mutate()

c. tidyr() - spread(), gather()

d. readr() - read_delim(), write_rds()


#R Basics
##Code Edit 1
The exclamation mark from the original object name was removed 
```{r}
My_data.name___is.too00ooLong <- c(1,2,3)
```

##Code Edit 2
The "it" was missing a closing parenthesis. The c defining the struct was capitalized.

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```

##Vector Storage
The entire vector is stored as characters rather than numerics (this differed from my expectation that only 3 and 4 would be stored as characters).

```{r}
my_vector <- c(1,2, '3', '4',5)
my_vector
is.numeric(my_vector[1])
```


#Data Import/Export
##Rail Trail Import
Import Rail Trail.txt:
```{r import, message=FALSE}
holder <- 'C:/Users/BryanHee/OneDrive - stok LLC/Intro to Data Science/HW Assignments/Assignment 5/rail_trail.txt'
rail_trail <- read_delim(file = holder, '|')
rm(holder)
glimpse(rail_trail)
```

##Rail Trail Export
Export Rail Trail.rds:
```{r export}
write_rds(rail_trail, path = 'C:/Users/BryanHee/OneDrive - stok LLC/Intro to Data Science/HW Assignments/Assignment 5/rail_trail.rds')
read_rds('C:/Users/BryanHee/OneDrive - stok LLC/Intro to Data Science/HW Assignments/Assignment 5/rail_trail.rds') %>%
glimpse()
```


#Visualization
##Mrs. President Graph Critique
##FIX
```{r}
ggplot(diamonds) +
  geom_boxplot(mapping = aes(cut, carat, position = color)) +
  labs(x = 'CUT OF DIAMOND', y = 'CARAT OF DIAMOND') +
  coord_flip()
```


##Question 3
```{r}
ggplot(diamonds, mapping = aes(cut, carat, fill = color)) +
  geom_boxplot() +
  labs(x = 'CUT OF DIAMOND', y = 'CARAT OF DIAMOND') +
  coord_flip()
```

#Data Munging and Wrangling
##Table2
No table2 is not currently tidy because the type column includes two variables.
```{r}
spread(table2, key = 'type', value = 'count')
```

##Price per Carat
```{r Price per Carat, echo=FALSE}
diamonds <- mutate(diamonds, 
       price_per_carat = price / carat)
```


##FIX
```{r}
pricy_small_diamonds <- diamonds %>%
  filter(price>10000 & carat < 1.5) #%>%
#summarise(count(cut))

  
```


#EDA - Exploratory Data Analysis
##Question 1
```{r}
texas_housing <- txhousing
head(texas_housing)
tail(texas_housing)
```

##Question 2
##Question 3
##Question 4
##Question 5
##Question 6