---
title: "COMPSCIX 415.2 Homework 3"
author: "Bryan Hee"
date: "February 13, 2018"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

## Section 3.8.1
### Question 1
The problem with the plot is that there is a quite a bit of overlapping data. To better communicate the information, you could use position = "jitter" to remove the overplotting effects.
```{r question_3.8.1.1}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter") + 
  geom_count()
```

### Question 2
The parameters that control the amount of jittering within the geom_jitter() geom are width and height

### Question 3
The geom geom_count counts the number of observations at each location and then displays that information as a count to point area. The geom geom_jitter on the otherhand changes the observations' values slightly to better see the disagregated sample graphically. geom_count is more useful if you are trying to understand exactly how much overplotting is occuring and where the highest densities of overplotting occur, whereas geom_jitter is more useful in seeing the representational data sample.


### Question 4



## Section 3.9.1

## Section 4.4

## Section 5.2.4

## Section 5.4.1

