---
title: "homework_2_Hee_Bryan"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


```{r loadplots}
library(tidyverse)
```

EXERCISE 3.2.4

1. All i see is a blank canvas or axis-set. See code below:
```{r exercise_3.2.4.1}
ggplot(data = mpg)
```

2. There are 234 rows and 11 columns in the mpg dataset

3. The drv variable describes the drive type of the car. i.e. whether the car is front-wheel drive, rear-wheel drive, or 4-wheel drive

4. See scatterplot of hwy vs. cyl below:
```{r exercise_3.2.4.4}
ggplot(data = mpg) + geom_point(mapping = aes(x = cyl, y = hwy))
```

5. The graph class vs. drv is not very useful because we are graphing discrete 1-to-1 categorical data as a scatter plot. It doesn't allow an easy visual understanding of how many classes belong to each drive type.See graph below:
```{r exercise_3.2.4.5}
ggplot(data = mpg) + geom_point(mapping = aes(x = drv, y = class))
```

EXERCISE 3.3.1

1. The following code does not plot the mpg dataset hwy vs. disp in the color blue because the code "color = "blue"" is included in the aes() function rather than outside aes() but still inside geom_point()
```{r exercises_3.3.1}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

2. The categorical variables included in the dataset mpg, include: manufacturer, model, trans, drv, fl, and class. The continuous variables included are: displ, year, cyl, cty, and hwy

3. 

4.

5. 


```{r exercises_3.5.1.4and5}

```

```{r exercises_3.6.1}

```

```{r exercises_3.7.1.2}

```

What is a Data Scientist?
