---
title: "homework_2_Hee_Bryan"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


```{r loadplots}
library(tidyverse)
```

EXERCISE 3.2.4

1. All i see is a blank canvas or axis-set. See code below:

```{r exercise_3.2.4.1}
ggplot(data = mpg)
```

2. There are 234 rows and 11 columns in the mpg dataset

3. The drv variable describes the drive type of the car. i.e. whether the car is front-wheel drive, rear-wheel drive, or 4-wheel drive

4.
```{r exercise_3.2.4.4}
ggplot(data = mpg) + geom_point(mapping = aes(x = cyl, y = hwy))
```

5. The graph class vs. drv is not very useful because we are graphing discrete 1-to-1 categorical data as a scatter plot. It doesn't allow an easy visual understanding of how many classes belong to each drive type. 
```{r exercise_3.2.4.5}
ggplot(data = mpg) + geom_point(mapping = aes(x = drv, y = class))
```


```{r exercises_3.3.1}

```

```{r exercises_3.5.1.4and5}

```

```{r exercises_3.6.1}

```

```{r exercises_3.7.1.2}

```

What is a Data Scientist?
