---
title: "COMPSCIX 415.2 Homework 2"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

EXERCISE 3.2.4

1. All i see is a blank canvas or axis-set. See code below:
```{r exercise_3.2.4.1}
ggplot(data = mpg)
```

2. There are 234 rows and 11 columns in the mpg dataset

3. The drv variable describes the drive type of the car. i.e. whether the car is front-wheel drive, rear-wheel drive, or 4-wheel drive

4. See scatterplot of hwy vs. cyl below:
```{r exercise_3.2.4.4}
ggplot(data = mpg) + geom_point(mapping = aes(x = cyl, y = hwy))
```

5. The graph class vs. drv is not very useful because we are graphing discrete 1-to-1 categorical data as a scatter plot. It doesn't allow an easy visual understanding of how many classes belong to each drive type.See graph below:
```{r exercise_3.2.4.5}
ggplot(data = mpg) + geom_point(mapping = aes(x = drv, y = class))
```

EXERCISE 3.3.1

1. The following code does not plot the mpg dataset hwy vs. disp in the color blue because the code "color = "blue"" is included in the aes() function rather than outside aes() but still inside geom_point()
```{r exercises_3.3.1}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

2. The categorical variables included in the dataset mpg, include: manufacturer, model, trans, drv, fl, and class. The continuous variables included are: displ, year, cyl, cty, and hwy

3. 

4.

5. 


```{r exercises_3.5.1.4and5}

```

```{r exercises_3.6.1}

```

```{r exercises_3.7.1.2}

```

What is a Data Scientist?
